---
title: "Fractal Dimensions with fractaldim library"
---

## Data

```{r,message=FALSE,warning=FALSE,results='hide'}
library(readr)
# Load the BTC data
btc_1d = read_csv("btc_1d_data_2018_to_2025.csv")
btc_1d$Date = as.Date(btc_1d$`Open time`)
btc_1d = btc_1d[,c("Date","High")]

```

## Beyond Euclidean Geometry
In standard geometry, we deal with integer dimensions: a point (0D), a line (1D), and a plane (2D). However, natural phenomena—like the jagged path of a stock price or the silhouette of a mountain—often exist in the "fractional" space between these integers.

**Fractal Dimension ($D$)** measures how a pattern's detail changes with the scale at which it is measured. The most intuitive definition is the **Scaling Law**: 

If you reduce the measurement scale by a factor of $L$ and find $N$ number of self-similar pieces, the dimension is:
$$D = \frac{\log(N)}{\log(L)}$$





## Interpreting the Values
When analyzing a 1D time series, your calculated $D$ will typically fall between **1.0** and **2.0**. Here is how to interpret the "physics" of those numbers:

| Fractal Dimension ($D$) | Description | Visual Texture |
| :--- | :--- | :--- |
| **$D \approx 1.0$** | Differentiable | Smooth, predictable curves or straight lines. |
| **$1.2 < D < 1.3$** | Low Complexity | Gentle oscillations with minor noise. |
| **$D \approx 1.5$** | Brownian Motion | The "Random Walk." Unpredictable with no long-term memory. |
| **$D \to 2.0$** | Space-Filling | Extremely "jagged" and volatile; the line moves so much it nearly fills a 2D area. |





## The fractaldim Library in R
The `fractaldim` package is the preferred tool for high-performance estimation. It provides several algorithms to calculate $D$, as different methods have different sensitivities to noise and signal length.

### Key Estimation Methods

* **Box-Counting (`boxcount`):** The most famous method. It covers the signal with a grid of boxes of size $\epsilon$ and counts how many boxes are intercepted by the signal.

* **Variogram (`variogram`):** Analyzes the variance of increments. This is often the most robust method for signals that resemble stochastic processes.

* **Hall-Wood (`hallwood`):** A specialized estimator designed to reduce bias in small samples or very noisy data.



## Practical Implementation
To use this in your workflow, you generally pass a numeric vector to the `fd.estimate` function.

```{r}
library(fractaldim)

# Estimate the dimension using multiple methods
# It is best practice to compare a few methods to ensure consistency
estimate.b <- fd.estimate(btc_1d$High, methods = "boxcount")
estimate.v <- fd.estimate(btc_1d$High, methods = "variogram")
```

```{r}
# Access the numerical results
cat("Estimated D (Box-count):", estimate.b$fd)
cat("Estimated D (Variogram):", estimate.v$fd)
```